---
title: "EDA"
output: html_notebook
---


```{r}
PlayList <- read.csv("PlayList.csv")
InjuryRecord <- read.csv("InjuryRecord.csv")
library(tidyverse)
library(ggplot2)
```

```{r}
glimpse(PlayList)
summary(PlayList)
colnames(PlayList)
```
This is the detailed summary of the various variables within the data set and their classification type. Below is a more detailed description of each variable...

PlayerKey: Player Identification number to be used instead of name

PlayerGame: The number of games the player had played in till the injury

PlayType: Description of the various types of plays that are possible in the NFL. These include passing downs as well as rushing downs.

GameID: Game identification number

StadiumType: Description of the structure of the stadium the player played at.

PlayerGamePlay: The number of games the player has played.

PlayKey: Play identification number to reference specific type of play that occured.

FieldType: Surface of the field ranging from synthetic turf to natural grass.

Position: Identifies the roster position of the player.

RosterPosition: Identifies the roster position of the player.

Temperature: Indicates the temperature at the time of play during the specific game.

PositionGroup: Identifies the roster position of the player.

PlayerDay: The days since the start of the season for the specific player.

Weather: Descriptions of the weather conditions during the specific game and time of play.

```{r}
summary(as.factor(PlayList$RosterPosition))
summary(as.factor(PlayList$StadiumType))
summary(as.factor(PlayList$FieldType))
summary(as.factor(PlayList$Weather))
summary(as.factor(PlayList$PlayType))
summary(as.factor(PlayList$Position))
summary(as.factor(PlayList$PositionGroup))
```
Shows the number of each category that exists within the PlayList dataset.

```{r}
glimpse(InjuryRecord)
summary(InjuryRecord)
colnames(InjuryRecord)
```

```{r}
summary(as.factor(InjuryRecord$BodyPart))
summary(as.factor(InjuryRecord$Surface))
```
Shows that the most lower body injuries occured at the Knee and most lower body injuries occured on synthetic field surfaces.

```{r}
#bar chart
ggplot(data = nflData) +
  geom_bar(mapping = aes(x = RosterPosition))
```
Shows that linebackers contracted the most injuries out of any position group. This makes sense given the rough nature of the position.

```{r}
# histogram
nflData %>% #pipeline expression
  ggplot() +
  geom_histogram(mapping = aes(x = PlayerGamePlay), binwidth = 0.1)
```
This shows that the most number of injuries occured during the start of the season and declined at a gradual rate as the season progressed.

```{r}
nflData %>% 
  ggplot(mapping = aes(x = PlayerGamePlay, colour = PlayType)) +
  geom_freqpoly(binwidth = 0.1)
```
This shows that most injuries occured during pass plays compared to rush plays which could be a suggestion of poor pass protection as well as pass interference calls that could have contributed to a player's injury.

```{r}
# boxplot (one categorical + one numerical variable)
nflData %>%
  ggplot(mapping = aes(x = Position, y = PlayerGame)) +
  geom_boxplot()
```
The higher injury rates occur at the defensive positions compared to offensive positions which may be a sign of dangerous game play on the defensive side of the ball. 

```{r}
# two categorical variables
nflData %>% 
  ggplot() +
  geom_count(mapping = aes(x = Position, y = PlayType))
```
Rush and Pass plays are equally balanced to show that the data contains instances of injury for both rushing and passing downs.


Now for the analysis of lower body injuries from the Injury Record data set.
```{r}
ggplot(data = InjuryRecord) +
  geom_bar(mapping = aes(x = BodyPart))
```
Evidently, the knee seems to be at high risk of injury according to injury records, whereas the heel seems to have the lowest risk of injury.

```{r}
ggplot(data = InjuryRecord) +
  geom_bar(mapping = aes(x = Surface))
```
More injuries occured on synthetic field surfaces.


